Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2017-12-22 PAPER.tex
made plots larger

\begin{figure}
	\centering
	\includegraphics[trim={1.5cm 0cm 3.5cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset. We model a mock foreground-only visibility with a sinusoid signal that varies smoothly in 
time and frequency. We model a mock EoR signal as a random Gaussian signal. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- 
foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$ mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare inverse covariance weighting 
where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former 
case, signal loss arises from using information from all eigenmodes of $\hat{\textbf{C}}$. Because $\hat{\textbf{C}}$ is 
empirically estimated, its eigenvalues differ from those of the true covariance and this error primarily affects the weaker (EoR-dominated) eigenmodes.
There is no signal loss when using only the first eigenmode, since we are not using information from the weaker eigenmodes 
which $\hat{\textbf{C}}$ does not describe accurately.}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ 
(green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one 
sinusoid. These empirically estimated covariance matrices have eigenspectra that are different from that of a true $\textbf{C}$, 
meaning that some eigenmodes (shapes in the data) may be down-weighted more significantly than they should be, producing 
signal loss. Specifically, down-weighting the weak eigenmodes, where the EoR signal is greater than the FG signal, leads to 
loss.}
	\label{fig:toy_sigloss2}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={1.5cm 0cm 3.5cm 0.1cm},clip,width=\columnwidth]{plots/toy_sigloss12.png}
	\includegraphics[trim={1.5cm 0cm 3.5cm 0.1cm},clip,width=\columnwidth]{plots/toy_sigloss13.png}
	\caption{The covariance matrices (top row) and inverse covariance-weighted data (bottom row) for FG only (left), EoR only 
(middle), and FG + EoR (right). Real parts are shown here.}
	\label{fig:toy_sigloss12}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={1.5cm 0cm 1.5cm 0.3cm},clip,height=0.29\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},clip,height=0.29\textwidth]{plots/toy_sigloss15.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, 
yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\hat{\textbf{C}}_{FG+EoR}$, 
in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green).}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0.3cm 0.2cm},clip,width=\columnwidth]{plots/toy_sigloss16.png}
	\caption{The convergence level, as defined in Equation \eqref{eq:converge}, of empirically estimated covariances of mock EoR signals with different numbers of independent samples. In red, the mock EoR signal is comprised entirely of independent samples. Subsequent colors show time-averaged signals. As the number of realizations increases, we see that the empirical covariances approach the true covariances. With more independent samples, the quicker an empirical covariance converges.}
	\label{fig:toy_sigloss16}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0.3cm 0.2cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the `fringe-rate filtered' (time-averaged) toy model simulation --- foregrounds only (blue), 
EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is 
computed from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a 
steepened, weakly characterized eigenspectrum from fewer independent modes in the data. The power spectrum 
estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes (the 
weakest, EoR-dominated modes).}
	\label{fig:toy_sigloss7}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0.4cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={1cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0.4cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={1cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our `fringe-rate filtered' (time-averaged) toy model simulation --- foregrounds only (blue), 
EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each minimize signal 
loss, including modeling the covariance matrix of EoR (upper left), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to 
it (upper right), using only the first three eigenmodes of $\hat{\textbf{C}}$ (lower left), and keeping only the diagonal elements of 
$\hat{\textbf{C}}$ (lower right).}
	\label{fig:toy_sigloss8}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate 
weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity 
matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
\hat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
computed for the `fringe-rate filtered' (time-averaged) toy model case.}
	\label{fig:toy_sigloss14}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation from bootstrapping as a function of the number of elements drawn per bootstrap when 
sampling with replacement. The star represents the standard deviation of $N_{boot}$ bootstraps, each created by drawing $1000$ 
elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be under-estimated if 
drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical 
prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error2.png}
	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a 
dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all 
elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought 
of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (gray) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} 
%The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
	\label{fig:toy_error2}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_bias1.png}
	\caption{Power spectrum estimates for a null jackknife test with the presence of a foreground systematic (blue), without 
the foreground systematic (red), and noise alone (green). Because the first null test is not consistent with noise, it suggests the 
presence of a systematic in either $\textbf{x}_{1}$ or $\textbf{x}_{2}$. Null tests of clean measurements should be consistent 
with thermal noise.}
	\label{fig:toy_bias1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_bias2.png}
	\caption{Power spectrum estimates for $\textbf{x}_{1}$ and $\textbf{x}_{2}$, two jackknives of the toy model. They suggest 
the presence of a systematic in $\textbf{x}_{1}$ only, illustrating how jackknives can be used to tease out excesses. Clean 
measurements should remain consistent despite the jackknife taken.}
	\label{fig:toy_bias2}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 4cm 0cm 5cm},width=\columnwidth]{plots/ant_layout_aspect.png}
	\caption{The PAPER-64 antenna layout. We use only the $30$ m East/West baselines for the revised analysis in this 
paper (i.e. the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}

>>>>

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset. We model a mock foreground-only visibility with a sinusoid signal that varies smoothly in 
time and frequency. We model a mock EoR signal as a random Gaussian signal. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- 
foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$ mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare inverse covariance weighting 
where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former 
case, signal loss arises from using information from all eigenmodes of $\hat{\textbf{C}}$. Because $\hat{\textbf{C}}$ is 
empirically estimated, its eigenvalues differ from those of the true covariance and this error primarily affects the weaker (EoR-dominated) eigenmodes.
There is no signal loss when using only the first eigenmode, since we are not using information from the weaker eigenmodes 
which $\hat{\textbf{C}}$ does not describe accurately.}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ 
(green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one 
sinusoid. These empirically estimated covariance matrices have eigenspectra that are different from that of a true $\textbf{C}$, 
meaning that some eigenmodes (shapes in the data) may be down-weighted more significantly than they should be, producing 
signal loss. Specifically, down-weighting the weak eigenmodes, where the EoR signal is greater than the FG signal, leads to 
loss.}
	\label{fig:toy_sigloss2}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss12.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss13.png}
	\caption{The covariance matrices (top row) and inverse covariance-weighted data (bottom row) for FG only (left), EoR only 
(middle), and FG + EoR (right). Real parts are shown here.}
	\label{fig:toy_sigloss12}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, 
yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\hat{\textbf{C}}_{FG+EoR}$, 
in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green).}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss16.png}
	\caption{The convergence level, as defined in Equation \eqref{eq:converge}, of empirically estimated covariances of mock EoR signals with different numbers of independent samples. In red, the mock EoR signal is comprised entirely of independent samples. Subsequent colors show time-averaged signals. As the number of realizations increases, we see that the empirical covariances approach the true covariances. With more independent samples, the quicker an empirical covariance converges.}
	\label{fig:toy_sigloss16}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the `fringe-rate filtered' (time-averaged) toy model simulation --- foregrounds only (blue), 
EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is 
computed from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a 
steepened, weakly characterized eigenspectrum from fewer independent modes in the data. The power spectrum 
estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes (the 
weakest, EoR-dominated modes).}
	\label{fig:toy_sigloss7}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our `fringe-rate filtered' (time-averaged) toy model simulation --- foregrounds only (blue), 
EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each minimize signal 
loss, including modeling the covariance matrix of EoR (upper left), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to 
it (upper right), using only the first three eigenmodes of $\hat{\textbf{C}}$ (lower left), and keeping only the diagonal elements of 
$\hat{\textbf{C}}$ (lower right).}
	\label{fig:toy_sigloss8}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate 
weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity 
matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
\hat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
computed for the `fringe-rate filtered' (time-averaged) toy model case.}
	\label{fig:toy_sigloss14}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation from bootstrapping as a function of the number of elements drawn per bootstrap when 
sampling with replacement. The star represents the standard deviation of $N_{boot}$ bootstraps, each created by drawing $1000$ 
elements (with replacement) from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be under-estimated if 
drawing more elements than there are independent samples in the data. The estimated errors match up with the theoretical 
prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_error2.png}
	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a 
dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all 
elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought 
of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (gray) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} 
%The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
	\label{fig:toy_error2}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias1.png}
	\caption{Power spectrum estimates for a null jackknife test with the presence of a foreground systematic (blue), without 
the foreground systematic (red), and noise alone (green). Because the first null test is not consistent with noise, it suggests the 
presence of a systematic in either $\textbf{x}_{1}$ or $\textbf{x}_{2}$. Null tests of clean measurements should be consistent 
with thermal noise.}
	\label{fig:toy_bias1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias2.png}
	\caption{Power spectrum estimates for $\textbf{x}_{1}$ and $\textbf{x}_{2}$, two jackknives of the toy model. They suggest 
the presence of a systematic in $\textbf{x}_{1}$ only, illustrating how jackknives can be used to tease out excesses. Clean 
measurements should remain consistent despite the jackknife taken.}
	\label{fig:toy_bias2}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/ant_layout_aspect.png}
	\caption{The PAPER-64 antenna layout. We use only the $30$ m East/West baselines for the revised analysis in this 
paper (i.e. the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}
